One-Step, Three-Factor Authentication in a Single Earpiece

### 2017-04-12 paper.tex
Reworking table 4, results text, and discussion text to interrogate factor presence

Mobile and wearable technologies' continuous proximity to the user's body provides favorable conditions for unobtrusively capturing biometrics for authentication. Many such uses have been proposed that embrace usability like touch-based interactions \cite{Tartz2015, Holz2015} and walking patterns \cite{Lu2014} using mobile phones, as well as identification via head movements and blinking in head-worn devices \cite{Rogers2015}. However, these typically draw only from the inherence factor. \cite{Chen2015} proposed an inherence and knowledge two-factor method for multi-touch mobile devices based on a user's unique finger tapping of a song, though it may be vulnerable to "shoulder surfing": imposters observing and mimicking the behavior. By comparison, passthoughts are largely invisible to attackers. Furthermore, this system also presents a solution to "rubber-hose attacks" (forceful coercion to divulge a password): a thought, which is secret (and thus changeable), but has a particular expression unique to an individual, the performance of which cannot be described (and thus cannot be coerced).

@@

We used a set of mental tasks based on findings in related work regarding the relative strengths of different tasks in authentication accuracy and usability as reported by participants. Furthermore, given the in-ear placement of the electrodes and therefore the proximity to the temporal lobes containing the auditory cortex, we tested several novel authentication tasks based specifically on aural imagery or stimuli. The 9 authentication tasks and their attributes are listed in Table \ref{tab:tasks}. Our strategy was to select tasks that captured a diversity across dimensions of external stimuli, involving a personal secret, eyes open or closed (due to known effects on EEG), and different types of mental imagery.

@@

We analyzed the EEG signals collected during the tasks using a support vector classifier (SVC). Since past work has shown that classification tasks in EEG-based brain-computer interfaces (BCI) are linear \cite{Garrett2003a}, we used XGBoost, a popular tool for logistic linear classification \cite{Chen2016}. Compared to other linear classifiers, XGBoost uses gradient boosting; in which an algorithm generates an ensemble (in this case, a decision tree) of weak linear classifiers that minimizes a given loss function. Gradient boosting gives better results in linear classification problems, without our needing to manually tune classifier hyperparameter.

@@

These results establish good performance in our original training strategy (between-participants, any task), in which we counted as negative examples recordings from the wrong participant performing any task. For comparison, we report two additional training strategies: one in which negative examples are the correct task recorded from the wrong participant (between participants, same tasks), and one in which negative examples are the wrong task recorded from the correct participant (within-participants). Positive examples were always correct participant, correct task.

\begin{table}[h]
\begin{center}
\begin{tabular}{lrr}
 \textbf{Analysis Scheme} & \textbf{FAR} & \textbf{FRR}\\
\hline
Between-participants, any task & 0.000074 & 0.004424\\
Between-participants, same task & 0.000724 & 0.001522\\
Within-participants & 0.002523 & 0.039702\\
\hline
\end{tabular}
\end{center}
\caption{Mean FAR and FRR for all participants and passthoughts across three analysis schemes.}
\label{tab:compare}
\end{table}

Overall, our between-participants training strategy achieves the lowest FAR (Table \ref{tab:compare}). Within-tasks FAR is ten times higher, and within-participants FAR is one hundred times higher, though for all scenarios still less than 1\%.
However, FRR is \textit{lower} in the within-tasks training strategy than in our original strategy. Within-participants again results in the highest FRR.

@@

To explore the scenario of an imposter attempting to gain access, we chose the case of the most vulnerable participant, P6, whose earpieces P1 had the lowest impedances while wearing. We collected data using the same data collection protocol, but had P1 refer to P6's report of chosen passsthoughts.
P1 performed each of P6's passthoughts (simulating an "inside imposter" from within the system). Following the same classification and preprocessing steps, we generated 200 samples per task for our imposters, using data from all left ear electrodes.

@@

\section{Discussion}

@@

This study was conducted with a small, relatively homogeneous sample of participants. In the case of this initial evaluation however, the homogeneity of our participant pool strengthens the reported results given that system is meant to distinguish between individuals. In order to establish the validity of this system for widespread real-world use we feel it is necessary to expand the size and diversity of participants in future studies. This sample size is comparable to that of other EEG passthoughts work \cite{Ashby2011, Marcel2007a, Poulos2002, Chuang2013b, curran2016passthoughts} and other custom-fit in-ear EEG research \cite{Kidmose2013a, Mikkelsen2015}. The fitting and manufacturing of custom-fit earpieces was the main limitation to increasing our sample size. This may very well pose a limitation in proliferation and adoption of such a technology, although recently there have been developments in at-home kits for creating one's own custom-fitted earpieces \cite{voix2015settable} that could help overcome this barrier.

Among our authentication analysis schemes, the between-participants analyses achieved lower FAR than within-participants, indicating the inherence factor (a person's brainwaves) is likely stronger than the knowledge factor (their choice of task), though the knowledge factor is not trivial. The \textit{breathe}, \textit{breathe-open}, and \textit{listen} tasks performed exceedingly well among participants. Classifiers overall distinguished the \textit{breathe} and \textit{listen} tasks even compared to the same tasks from other participants, implying that these tasks are expressed differently for each participant, i.e. that they have an inherence factor sufficient for authentication, even though they task do not involve an explicit secondary knowledge factor as in \textit{sport} or \textit{song}. We were able to achieve these results by generating feature vectors based on only 500ms (300 voltage readings across the three electrodes). This short timespan is somewhat surprising, given that tasks like \textit{song} presumably change over longer time periods.

Contrary to our expectations, relatively low impedances are still achievable by an imposter using other participants' custom-fit earpieces, despite the uniqueness of ear canal shapes between individuals \cite{Akkermans2005}. This could likely have been due to our use of conductive gel and warrants further exploration. This discovery is not detrimental, however, as the primary inherence factor of this system is drawn from the uniqueness of EEG and not ear canal shape. Classifiers appear to still resist spoofing attacks, indicating that task-related signals are unique to individuals. In terms of establishing a possession factor, the custom-fit earpieces could easily include a hardware keypair to sign authentication attempts in a three-factor, one-step scheme.

The powerful interactions between inherence and knowledge emerged in our spoofing attack. Although our target participant documented their chosen passthought, the spoofers found ambiguity in how these passthoughts could be expressed. For the \textit{face} task, the spoofers did not know the precise face the original participant had chosen. For the \textit{song} and \textit{song - open} tasks, though the song was known, the spoofers did not know what part of the song the original participant had imagined, or how it was imagined (humming, imagining a full performance, melody, vocals, etc). This experience sheds light on the highly individual nature of passthoughts, and provides a positive indication that there may be some intrinsic difficulty in spoofing attempts of passthoughts.

Finally, performance on Fp1 was not as high as performance in the ear, despite Fp1's popularity in past work on passthoughts \cite{Chuang2013b}. This could be explained by the greater number of electrodes in the ear in this analysis (compared to one on Fp1). Additionally, Fp1 is best poised to pick up on frontal lobe activity (e..g, concentration), but our tasks did not generally involve frontal lobe activity; in fact, a good number of them involved audio, which we would expect to be better observed from the auditory cortex near the ears. Future work should continue to investigate what sorts of mental tasks best lend themselves to in-ear recording.

>>>>

Mobile and wearable technologies' continuous proximity to the user's body provides favorable conditions for unobtrusively capturing biometrics for authentication. Many such uses have been proposed that embrace usability like touch-based interactions \cite{Tartz2015, Holz2015} and walking patterns \cite{Lu2014} using mobile phones, as well as identification via head movements and blinking in head-worn devices \cite{Rogers2015}. However, these typically draw only from the inherence factor. \cite{Chen2015} proposed an inherence and knowledge two-factor method for multi-touch mobile devices based on a user's unique finger tapping of a song, though it may be vulnerable to "shoulder surfing": imposters observing and mimicking the behavior. By comparison, passthoughts are largely invisible to attackers. Furthermore, this system also presents a possible solution to "rubber-hose attacks" (forceful coercion to divulge a password): a thought, which is secret (and thus changeable), but has a particular expression unique to an individual, the specific performance of which cannot be described (and thus cannot be coerced).

@@

We used a set of mental tasks based on findings in related work regarding the relative strengths of different tasks in authentication accuracy and usability as reported by participants. \cite{Chuang2013b, curran2016passthoughts} Furthermore, given the in-ear placement of the electrodes and therefore the proximity to the temporal lobes containing the auditory cortex, we tested several novel authentication tasks based specifically on aural imagery or stimuli. The 9 authentication tasks and their attributes are listed in Table \ref{tab:tasks}. Our strategy was to select tasks that captured a diversity across dimensions of external stimuli, involving a personal secret, eyes open or closed (due to known effects on EEG), and different types of mental imagery.

@@

We analyzed the EEG signals collected during the tasks using a support vector classifier (SVC). Since past work has shown that classification tasks in EEG-based brain-computer interfaces (BCI) are linear \cite{Garrett2003a}, we used XGBoost, a popular tool for logistic linear classification \cite{Chen2016}. Compared to other linear classifiers, XGBoost uses gradient boosting in which an algorithm generates an ensemble (in this case, a decision tree) of weak linear classifiers that minimizes a given loss function. Gradient boosting gives better results in linear classification problems without manually tuning hyperparameters.

@@

These results establish good performance in our default training strategy, in which we count as negative examples recordings from the wrong participant performing any task. For comparison, we report three other analyses with differing negative examples: the correct task recorded from the wrong participant (inherence only), the wrong task recorded from the correct participant (knowledge only), and a combination of these two. Positive examples were always correct participant, correct task.

\begin{table}[h]
\begin{center}
\begin{tabular}{llrrr}
 \textbf{+ve Ex.'s} & \textbf{-ve Ex.'s} & \textbf{Factors} & \textbf{FAR} & \textbf{FRR} \\
\hline
$P_c, T_c$ & $P_i, T_*$ & I & 0.000074 & 0.004424\\
$P_c, T_c$ & $P_i, T_c$ & I & 0.000724 & 0.001522\\
$P_c, T_c$ & $P_c, T_i$ & K & 0.002523 & 0.039702\\
$P_c, T_c$ & $P_i, T_* + P_c, T_i$ & I \& K & 0.000186 & 0.052565\\
\hline
\end{tabular}
\end{center}
\caption{Four analyses training classifiers on different negative examples with resulting mean FAR and FRR across all participants and tasks. $P_c$ indicates correct participant, $P_i$ incorrect participant, $T_c$ correct task, $T_i$ incorrect task, and $T_*$ any task. I indicates inherence factor and K knowledge factor was interrogated.}
\label{tab:compare}
\end{table}

Overall, our default training strategy achieves the lowest FAR (Table \ref{tab:compare}). Within-task FAR is ten times higher, and within-participants FAR is one hundred times higher, though for all scenarios still less than 1\%.
However, FRR is \textit{lower} in the within-tasks training strategy than the default. The highest FRR is in the combined negative example case, though FAR remains low.

@@

To explore the scenario of an imposter attempting to gain access, we chose the case of the most vulnerable participant, P6, whose earpieces P1 had the lowest impedances while wearing. We collected data using the same data collection protocol, but had P1 refer to P6's report of chosen passsthoughts.
P1 performed each of P6's passthoughts (simulating an "inside imposter" from within the system). Following the same analysis steps, we generated 200 samples per task for our imposters, using data from all left ear electrodes.

@@

\section{Discussion \& Limitations}

@@

The \textit{breathe}, \textit{breathe-open}, and \textit{listen} tasks performed exceedingly well among participants, despite not having an explicit secondary knowledge factor as in \textit{sport} or \textit{song}. We were able to achieve these results by generating feature vectors based on only 500 ms (300 voltage readings across the three electrodes), which is somewhat surprising given that tasks like \textit{song} presumably change over longer periods.

Our multiple classifier training strategies establish the presence of both knowledge and inherence factors as schemes isolating each achieve low FARs. The inherence factor performs strongest, while the knowledge factor contributes to a sizable increase in FRR. Interestingly, by training classifiers to reject both wrong participants and wrong tasks from the correct participant, thus incorporating both the knowledge and inherence factors, the FRR is highest suggesting multi-factor authentication's added security may come at a small cost to convenience. As preliminary evidence of a possession factor in the current prototypes, we had P1 use their best task (listen) with P6's earpieces but tested on P1's classifier, with a resultant FAR of 0. The custom-fit earpieces could easily include a hardware keypair to sign authentication attempts presenting a more explicit possession factor.

The powerful interactions between inherence and knowledge emerged in our spoofing attacks. Although our target participant documented their chosen passthought, the spoofers found ambiguity in how these passthoughts could be expressed. For the \textit{face} task, the spoofers did not know the precise face the original participant had chosen. For the \textit{song} tasks, though the song was known, the spoofers did not know what part of the song the original participant had imagined, or how it was imagined. This experience sheds light on passthoughts' highly individual nature and suggests there may be intrinsic difficulty in spoofing attempts.

Contrary to our expectations, relatively low impedances were still achievable by an imposter using other participants' custom-fit earpieces, despite the uniqueness of ear canal shapes between individuals \cite{Akkermans2005}. This may have been due to our use of conductive gel and warrants further exploration. This discovery is not detrimental, however, as the primary inherence factor of this system is drawn from the uniqueness of EEG and not ear canal shape.

Finally, performance on Fp1 was not as high as performance in the ear, despite Fp1's popularity in past work on passthoughts \cite{Chuang2013b}. This could be explained by the greater number of electrodes in the ear in this analysis (compared to one on Fp1). Additionally, Fp1 is best poised to pick up on frontal lobe activity (e.g, concentration), that our tasks did not generally involve; in fact, several involved audio, which we would expect to be better observed from the auditory cortex near the ears. Future work should continue to investigate what sorts of mental tasks best lend themselves to in-ear recording.

The real-world implications of this study are limited by the small, relatively homogeneous sample of participants. In the case of this initial evaluation however, the homogeneity of our participant pool strengthens the reported results given that system is meant to distinguish between individuals. In order to establish the validity of this system for widespread real-world use we feel it is necessary to expand the size and diversity of participants in future studies. This sample size is comparable to that of other EEG passthoughts work \cite{Ashby2011, Marcel2007a, Poulos2002, Chuang2013b, curran2016passthoughts} and other custom-fit in-ear EEG research \cite{Kidmose2013a, Mikkelsen2015}. The fitting and manufacturing of custom-fit earpieces was the main limitation to increasing our sample size. This may very well pose a limitation in proliferation and adoption of such a technology, although recently there have been developments in at-home kits for creating one's own custom-fitted earpieces \cite{voix2015settable} that could help overcome this barrier.
